---
title: "Influenzanet Central database"
format: 
  html:
    code-fold: false
    df-print: kable

---

# Central Database (individual data)

This page show various inputs about central database with individual data 

```{r include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo=FALSE)
library(dplyr)
library(RSQLite)
library(DBI)
library(tidyr)
library(dplyr)
library(ggplot2)
source("conf.R")

init.path('db')

db = DBI::dbConnect(RSQLite::SQLite(), local_data_path('imports.db'))
#
imports = dbGetQuery(db, "select * from imports where time > date('now', '-8 days') order by time desc")

imports = imports |> 
              mutate(
                      hour=trunc(as.POSIXct(time), units="hours"), 
                      label=paste(rows, " (",dir,")")
              ) |>
              arrange(desc(hour), country)

stats_file = my.path('stats.json')

stats_time = file.mtime(stats_file)
stats = jsonlite::read_json(stats_file)

stats = bind_rows(stats)

start_date = monday_of_date(Sys.Date()) - 3 * 7
season = get_current_season()

rows = NULL

for(survey in c('intake', 'weekly', 'vaccination')) {
  r = survey_load_results(survey, cols=c("timestamp", "country"), date = list(min=start_date), season=season)
  r$id = NULL
  r = r |> arrange(timestamp) |> mutate(yw=iso_yearweek(timestamp))
  r$yw = iso_yearweek(r$timestamp)
  
  rr = r %>% group_by(yw, country) %>% summarize(n=n(), participants=n_distinct(person_id), .groups = "drop")
  rr$survey = survey
  rows = bind_rows(rows, rr)
}

rows = rows |> arrange(yw, country) |> mutate(label=paste0(n, " (", participants, ")"))

```

```{r}

rr = tidyr::pivot_wider(rows |> select(country, survey, yw, label), id_cols=c('country', 'survey'), names_from = yw, values_from = label)

rr

```

### Database imports from Influenzanet uploads (new platform exports)

```{r, echo=FALSE, warning=FALSE}
if(nrow(imports) > 0) {
ggplot(imports, aes(x=as.POSIXct(hour), y=rows, fill=country)) + 
  geom_bar(stat="identity") + 
  facet_grid(rows=vars(table), scales="free_y") +
   labs(x="Date")
} else {
  print("No imports data")
}
```

```{r, echo=FALSE, warning=FALSE}
if(nrow(imports) > 0) {
  tidyr::pivot_wider(imports, id_cols=c('country', 'hour'), names_from = "table", values_from = "label")
} else {
  print("No imports data")
}
```

# Data in tables by season

Data updated `r stats_time`

- 'pollster' = Imported table from the platform (only legacy Influenzanet platform)
- 'epidb' = Cumulated table from all imports (New platform & old)

```{r, echo=FALSE, warning=FALSE, fig.width=12, fig.height=8}
ggplot(stats, aes(x=season, y=count, fill=type)) + 
  facet_grid(rows=vars(table), cols=vars(country), scales="free_y") + 
  geom_bar(stat="identity") +
  theme(axis.text.x=element_text(angle=90))
```


