---
title: "Incidence data"
toc: true
format: 
  html:
    page-layout: full
    df-print: kable
---

```{r, include=FALSE}

knitr::opts_chunk$set(echo=FALSE)
library(tidyr)
library(dplyr)
library(ggplot2)

source("conf.R")
season = get_current_season()
#season = 2019

censore.threshold = .35

season_title = paste0(season, "-", season + 1)

init.path('ecdc/indicator')

bundles = readRDS(my.path("bundles.rds"))

plot_inc_superpose = function(ii, title, date_breaks="2 weeks") {
  ggplot(ii, aes(x=monday_of_week(yw), y=incidence * 1000, color=country)) + 
  geom_line() +
  scale_x_date(date_breaks = date_breaks) +
  labs(x="Week (monday date)", y="Incidence rate (/1000 participant)", title=paste("Incidence rate for ", title), caption = ifn.copyright()) +
  theme(legend.position = "top", axis.text.x = element_text(angle=90)) 
}

plot_count = function(ii, title) {
p = tidyr::pivot_longer(ii %>% select(yw, country, count, part), c('count','part') )
p$name = factor(p$name, c('part','count'))
ggplot(p, aes(x=monday_of_week(yw), y=value, fill=name)) + 
  geom_bar(stat="identity") +
  facet_grid(rows=vars(country, name), scales="free_y") +
  scale_fill_discrete(labels=c("part"="Active participants", "count"="Participant with syndrome")) +
  labs(x="Week (monday date)", y="Number of participants", title=paste("Participants count used to compute incidence for", title), caption = ifn.copyright()) +
  theme(legend.position = "top", axis.text.x = element_text(angle=90)) 
}

#' Compute weighted moving average, fixed window 2
#' x = vector of values
#' n = vector of weights
conv = function(x, n) {
  sapply(seq_along(x), function(i) {
      if(i == 1) {
        return(x[i])
      }
      r = min(i, 2) - 1
      w = n[(i-r):i]
      sum(x[(i-r):i] * w) / sum(w)
  })
}

#' Compute threshold value to censor data below a given number of maximum active participants
#' Max Active participants is estimated with the mean of max of 2 previous season, weighted by number of weeks 
#' So in the starting of a season, the previous seasons (will all weeks) count a little more on mean.
#' Limit to censor a week if active participant of a week  if below a proportion (threshold) of the estimated max.
compute_censoring = function(ii, threshold=.35) {
  max.active = ii %>% group_by(country, season) %>% summarize(max_active=max(part), n=n(), .groups="drop")
  max.active = max.active %>% 
    arrange(season) %>% 
    group_by(country) %>% 
    mutate(
      max_active.mav=conv(max_active,n)
    )
  ii = left_join(ii, max.active, by=c("country","season"))
  ii = ii %>% mutate(active_limit=.35 * max_active.mav, censored=!is.na(active_limit) & part < active_limit)
  ii
}



```

```{r}
cur.inc = bundles$incidence %>% filter(season == !!season)

```

To adjust incidence scale, some points are censored if active participant of a week is below `r censore.threshold * 100` % of the maximum participant of a season (with a mean on 2 previous seasons to be more accurate on the start of the season)

# ILI Ecdc Definition

## Current season (`r season_title`)

```{r}
ii = cur.inc %>% filter(syndrome == "ili.ecdc")

method.country = ii %>% group_by(country, method) %>% count() %>% ungroup()
method.country = method.country %>% group_by(country) %>% mutate(n_country=n())

ii = left_join(ii, bundles$active[, c('yw','country','method', 'active')], by=c('yw','country','method'))
ii$part[is.na(ii$part)] = ii$active[is.na(ii$part)]

ili.countries = n_distinct(ii$country)

ii = compute_censoring(ii)

```

### Incidence 

```{r}
h = ili.countries * .5
```

```{r, fig.width=12, fig.height=h}

ggplot(ii %>% filter(!censored), aes(x=monday_of_week(yw), y=country, fill=1000*incidence)) + 
  geom_tile(color="gray") +
  scale_fill_viridis_c(option = "plasma")

```

#### Superposed incidence for all countries

```{r, fig.width=12, fig.height=10}
plot_inc_superpose(ii, title="ILI (ECDC)")
```

### Count & Active participants

```{r}
h = ili.countries * 1.5
```

```{r, fig.width=12, fig.height=h, warning=F}
plot_count(ii, title="ILI (ECDC)")
```

## All seasons

```{r}
h = ili.countries * 1.4
```

```{r, fig.width=12, fig.height=h, warning=FALSE }
ii = bundles$incidence %>% filter(syndrome == "ili.ecdc")
ii = left_join(ii, bundles$active[, c('yw','country','method', 'active')], by=c('yw','country','method'))
ii$part[is.na(ii$part)] = ii$active[is.na(ii$part)]

ii = compute_censoring(ii)

ii = ii %>% filter(!(season == 2023 & country == "UK"))

plot_inc_superpose(ii %>% filter(!censored), title="ILI (ECDC)", date_breaks = "1 month") +
  geom_vline(data = ii %>% filter(censored), aes(xintercept = monday_of_week(yw)), colour="#D4D4D4", alpha=70) +
  facet_grid(cols=vars(season), rows=vars(country), scales="free_x") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90))

```

Gray vertical lines indicates that a week is not shown, because the number of active participants are below the threshold (`r censore.threshold * 100` % of max active participants)