---
title: "Who Covid19 Vaccination & Influenzanet"
author: "Clément Turbelin"
date: "16/12/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(ggplot2)

source("conf.R")

theme_set(theme_minimal())

init.path("who")

r = readRDS(my.path("who_vacc.rds"))

collect = r$data

extract_syndrome = function(data, syndrome, active.column) {
  act = sym(active.column)
  ii = data %>% filter(syndrome == !!syndrome) 
  ii = ii %>% group_by(country, season) %>% 
        mutate(
          max_active=max(!!act),
          censor=!!act >= .35 * max_active
        ) %>% ungroup()
  ii
}


labels.status = c('FALSE'='Unvaccinated', 'TRUE'='Vaccinated')

gg_week = function(data) {
  ggplot(data, aes(x=monday_of_week(yw)) ) +
    labs(x="Week (monday)",  caption=ifn.copyright())
}

plot_count_weekly = function(ii, syndrome, active.column=NULL) {
  g = gg_week(ii) + aes(color=status)
  
  if(!is.null(active.column)) {
    g = g +   geom_bar(aes(y=!!sym(active.column), fill="lightgray", stat="identity")) 
  }
  
  g = g + geom_line(aes(y=count, linetype="count") ) 
    
  
  g = g +
    facet_grid(rows=vars(status.type), cols=vars(country,season), scales="free_x") +
    scale_linetype_discrete(c('count'='solid','active'='dashed'), labels=c('count'='Syndrome count','active'='Active participant')) +
    theme_with("x_vertical") +
    guides(linetype=guide_legend("Counts"), color=guide_legend("Covid Vaccine status"))
  g
}

plot_prop_weekly = function(ii, syndrome, active.column) {
  gg_week(ii) + 
    geom_line(aes(y=1000 * count / !!sym(active.column), color=status, linetype=status.type) ) + 
    facet_grid(cols=vars(season), rows=vars(country), scales="free_x") +
    labs(
      subtitle = paste0("Syndrome ", syndrome, ", Active = w0"), 
      y="Incidence rate / 1000 participants"
    ) +
    scale_color_discrete(labels=c('FALSE'='Uncomplete', 'TRUE'='Complete')) +
    theme_with("x_vertical") +
    guides( color=guide_legend("Covid Vaccine status"), linetype=guide_legend("Vaccine state Type"))
}


```

# Cohort approach

```{r echo=FALSE}
inc.cohort = collect$inc %>% filter(status.type == "status.cohort" & !is.na(status))
```

- Vaccinated = If reported as fully vaccinated until an endpoint date (Sept of IT, UK, CH), 15 Nov for FR
- Non vaccinated = If reported as non vaccinated until now (biais ?)

```{r echo=FALSE, fig.width=10}

ss = collect$cohort %>%
      mutate(
        time=if_else(status.cohort, vaccinated.time, unvaccinated.time),
        date=as.Date(trunc(time, "days"))
      ) %>%
      group_by(date, status.cohort, country) %>%
      summarise(
        count=n(),
        .groups = "drop"
      ) %>%
      arrange(date) %>%
      group_by(status.cohort, country) %>%
      mutate(cum_count=cumsum(count)) %>%
      ungroup()
    
ggplot(ss, aes(x=date, y=cum_count, color=status.cohort)) + 
  geom_line() +
  facet_grid(rows=vars(country), scales="free_y") +
  ylim(0, NA) +
  scale_color_discrete(labels=labels.status) +
    
  labs(
    title="Count for vaccinated & Non Vaccinated populations"
  )

```

## Participants by week

Count of reported participants considered as active (denominator for incidence) for syndrome and PCR
For PCR, participants with reported PCR test done (regardless result availability) is counted as active (Note that in the platforms only participants with symptoms are asked to report if a PCR test is done). This infers that a non reported result is considered as negative.

```{r echo=FALSE, fig.width=10}

gg_week(inc.cohort) + 
  geom_line(aes(y=active.pcr.positive, color="active.pcr")) +
  geom_line(aes(y=active, color="active")) +
  facet_grid(rows=vars(country), cols=vars(status), scales="free_y", labeller = labeller(status=labels.status)) +
  scale_linetype_discrete(labels=labels.status) +
  scale_color_discrete(labels=c('active'='Accounting symptoms', 'active.pcr'='Accounting PCR test')) +
  guides(linetype=guide_legend("Syndrome"), color=guide_legend("Population type")) +
  labs(title="Count of weekly active participants by type & vaccination status")

```

## Syndromes 

Count of reported syndrome (color) by population (line type). Count of participants with at least one report matching the given syndrome definition.
If participant reported several times the syndrome during the considered week, he is only counted once.
"pcr.positive" is the non clinical syndrome definition defined as "reported a positive PCR test and any symptom" (since only participant with symptom can report a test result).

```{r echo=FALSE, fig.width=10}

gg_week(inc.cohort) + 
  geom_line(aes(y=count, color=syndrome, linetype=status)) +
  facet_grid(rows=vars(country), scales="free_y") +
  scale_linetype_discrete(labels=labels.status) +
  guides(linetype=guide_legend("Covid Vaccine status"), color=guide_legend("Syndrome")) +
  labs(title="Count of reported syndrome by population (by vaccinated status) ")

```

## Participant with PCR positive test

Count of participants (gray bars give the count of participants having a test for the same population), line the count of positive reported tests.

```{r echo=FALSE, fig.width=10, height=6}

ii = extract_syndrome(inc.cohort, "pcr.positive", "active.pcr.positive")

gg_week(ii) +
  geom_bar(aes(y=active.pcr.positive), fill="lightgrey", stat="identity") +
  geom_line(aes(y=count)) +
   labs(
    title="Weekly counts of pcr.positive by vaccination status (cohort)"
  ) +
  scale_color_discrete(labels=labels.status) +
  facet_grid(rows=vars(country), cols=vars(status), scales="free_y", labeller=labeller(status=labels.status))

```

Weekly proportion of positive test by population

```{r echo=FALSE, fig.width=10, height=6}

gg_week(ii) +
  geom_line(aes(y=1000 * count / active.pcr.positive, color=status)) +
   labs(
    title="Weekly proportion of `pcr.positive` by vaccination status (cohort approach)"
  ) +
  scale_color_discrete(labels=labels.status) +
  facet_grid(rows=vars(country))

```

# Using syndrome *covid.ecdc*

Count of participaants with syndrome (line) and count of active participant for each week (gray bars) - Active is using w0 method (=having a report on the considered week)

```{r echo=FALSE, fig.width=10}

ii = extract_syndrome(inc.cohort, "covid.ecdc", "active")

gg_week(ii) +
  geom_bar(aes(y=active), fill="lightgrey", stat="identity") +
  geom_line(aes(y=count)) +
   labs(
    title="Weekly counts of covid.ecdc by vaccination status (cohort)"
  ) +
  facet_grid(rows=vars(country), cols=vars(status), scales="free_y", labeller=labeller(status=labels.status))

```

Weekly proportion for each population vaccinated/unvaccinated (vaccinated status by cohort approach)

```{r echo=FALSE, fig.width=10, height=6}

gg_week(ii) +
  geom_line(aes(y=1000 * count / active, color=status)) +
   labs(
    title="Weekly counts of pcr.positive by vaccination status (cohort)"
  ) +
  scale_color_discrete(labels=labels.status) +
  facet_grid(rows=vars(country))

```

# Comparing other vaccination status approaches

- status.cohort = Vaccination status defined by cohort (with report date endpoint)
- status.last = Vaccination status is fixed with the last vaccination report available (regardless date, even in future from the computed week)
- status.timely = Vaccination status is computed using the last vaccination report available at each time point (even in future)

The ".5m" version include a limite of the vaccination state, not considered as vaccinated after 5 month after the last shot

## Clinical Syndrome covid.ecdc

Weekly count of participant with a report for "covid.ecdc" by vaccination status 

````{r, echo=FALSE, fig.width=10}

ii = collect$inc %>% filter(syndrome == "covid.ecdc")

gg_week(ii) +
  geom_line(aes(y=count, color=status)) +
  facet_grid(rows=vars(country), cols=vars(status.type), scales="free_y") +
  scale_color_discrete(labels=labels.status)

```

Counts for active (solid line), and report with syndrome  (dashed)

````{r, echo=FALSE, fig.width=10}

gg_week(ii) +
  geom_line(aes(y=count, color=status, linetype="count")) +
  geom_line(aes(y=active.pcr.positive, color=status, linetype="active")) +
  facet_grid(rows=vars(country), cols=vars(status.type), scales="free_y") +
  scale_linetype_manual(labels=c('active'='Active (PCR Done)', 'count'='PCR+'), values=c('active'='solid','count'='dashed')) +
  scale_color_discrete(labels=labels.status)

```

## Using PCR test 

Weekly count of participant with a report for "pcr.positive" by vaccination status   

````{r, echo=FALSE, fig.width=10}

ii = collect$inc %>% filter(syndrome == "pcr.positive")

gg_week(ii) +
  geom_line(aes(y=count, color=status)) +
  facet_grid(rows=vars(country), cols=vars(status.type), scales="free_y") +
  scale_color_discrete(labels=labels.status)


```

Counts for active (solid line), and with positive test report (dashed)


````{r, echo=FALSE, fig.width=10}

gg_week(ii) +
  geom_line(aes(y=count, color=status, linetype="count")) +
  geom_line(aes(y=active.pcr.positive, color=status, linetype="active")) +
  facet_grid(rows=vars(country), cols=vars(status.type), scales="free_y") +
  scale_linetype_manual(labels=c('active'='Active (PCR Done)', 'count'='PCR+'), values=c('active'='solid','count'='dashed')) +
  scale_color_discrete(labels=labels.status)

```


